Comparison of Catalogs

### 2024-07-25 compare/compare.qmd
fixed sigma_clip

```{python}
#| echo: false
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support
#quantity_support()
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import seaborn as sns
import pandas as pd
import glob
import os
plt.style.use('ggplot')

@@

```{python}
def compare_data(x, y, unc=None, sigma = False):
    """
    Performs a linear comparison between two datasets.

@@

    Parameters:
    - x (array-like): The independent variable data.
    - y (array-like): The dependent variable data.
    - unc (array-like with units, optional): The uncertainties associated with the independent variable data. Default is None.

    Returns:
    tuple: A tuple containing the following elements:
        - slope (float): The slope of the fitted linear model.
        - intercept (float): The intercept of the fitted linear model.
        - r2 (float): The R-squared value, indicating the proportion of variance explained by the linear model.
        - corr (float): The Pearson correlation coefficient, measuring the linear correlation between x and y.
    """
    try:
        x_data = np.ma.array(x.value, mask=np.isnan(x.value))
        y_data = np.ma.array(y.value, mask=np.isnan(y.value))
        if unc is None:
            w = None
        else:
            unc = np.ma.array(unc.value, mask=np.isnan(x.value))
            w = 1 / unc
        # initialize a linear fitter
        fit = fitting.LinearLSQFitter()

@@

        # fit the data with the fitter
        # check if sigma
        if sigma is False:
            fitted_line = fit(line_init, x_data, y_data, weights=w)
            outliers = None
        else:
            or_fit = fitting.FittingWithOutlierRemoval(fit, sigma_clip, niter=3, sigma=3)
            fitted_line, outliers = or_fit(line_init, x, y, weights=w)
        slope = fitted_line.slope.value
        intercept = fitted_line.intercept.value

@@

        # Remove NaN values
        mask = ~np.isnan(y_data) & ~np.isnan(y_pred)
        y_data_clean = y_data[mask]
        y_pred_clean = y_pred[mask]
        mask = ~np.isnan(x_data)
        x_data_clean = x_data[mask]

@@

        return slope, intercept, r2, corr, outliers
    except Exception:
        return 0, 0, 0, 0, None
```

@@

    # Perform linear comparison between the datasets
    if yerr is not None:
        unc = yerr
    elif yerr is None and xerr is not None:
        unc = xerr
    else:
        unc = None
    c, m, r2, corr, outliers = compare_data(x, y, unc, sigma)

    # Plot the data with error bars
    # errorbar -> needs np.ma -> no units
    if outliers is not None:
        filtered_data = np.ma.masked_array(y_data, mask = outlier)
        plt.scatter(x, filtered_data,  label="Fitted Data", c = "blue", fillstyle = "filled")
        plt.errorbar(x, y, yerr=yunc, fillstyle="filled", c = "black", label="Clipped Data", fmt = "o")
    else:
        plt.errorbar(x_data, y_data, xerr = xerr_d, yerr = yerr_d, alpha = 0.3, c = "blue", label = "Data", fmt = ".")

@@

```
## Luminosities

```{python}
#| echo: false
logKLum_corr = round(compare_data(dt["logKLum"], dt["logL_K"], sigma = True)[3], 3)
```

@@

```{python}
scatter_plot(dt["logKLum"], dt["logL_K"])
```

@@

```{python}
#| echo: false
mag_B_corr = round(compare_data(dt["mag_B"], dt["BT"],unc = dt["E_BT"], sigma = False)[3], 3)
Kmag_corr = round(compare_data(dt["Kmag"], dt["K"], unc = dt["E_K"])[3], 3)
```

@@

```{python}
scatter_plot(dt["mag_B"], dt["BT"], yerr = dt["E_BT"], sigma = True)
scatter_plot(dt["Kmag"], dt["K"], yerr = dt["E_K"])
```

@@

# Morphology

>>>>

```{python}
#| echo: false
import numpy as np
import matplotlib.pyplot as plt
from sklearn.metrics import r2_score
from astropy.io import ascii
from astropy.coordinates import SkyCoord
from astropy.table import Table, join, QTable
import astropy.units as u
from astropy.visualization import quantity_support
#quantity_support()
from astropy.stats import sigma_clip
from astropy.modeling import models, fitting
from scipy.stats import pearsonr
import seaborn as sns
import pandas as pd
import glob
import os
plt.style.use('ggplot')

@@

```{python}
def compare_data(x, y, sigma = False):
    """
    Performs a linear comparison between two datasets.

@@

    Parameters:
    - x (array-like): The independent variable data.
    - y (array-like): The dependent variable data.
    - unc (array-like with units, optional): The uncertainties associated with the a variable data. Default is None.

    Returns:
    tuple: A tuple containing the following elements:
        - slope (float): The slope of the fitted linear model.
        - intercept (float): The intercept of the fitted linear model.
        - r2 (float): The R-squared value, indicating the proportion of variance explained by the linear model.
        - corr (float): The Pearson correlation coefficient, measuring the linear correlation between x and y.
    """
    try:
        x_data = np.ma.array(x.value, mask=np.isnan(x.value))
        y_data = np.ma.array(y.value, mask=np.isnan(y.value))
        # initialize a linear fitter
        fit = fitting.LinearLSQFitter()

@@

        # fit the data with the fitter
        # check if sigma
        if sigma is False:
            fitted_line = fit(line_init, x_data, y_data)
            outliers = None
        else:
            or_fit = fitting.FittingWithOutlierRemoval(fit, sigma_clip, niter=3, sigma=2)
            fitted_line, outliers = or_fit(line_init, x_data, y_data)
        slope = fitted_line.slope.value
        intercept = fitted_line.intercept.value

@@

        # Remove NaN values
        mask = ~np.isnan(y_data)
        if outliers is not None:
            mask = ~np.isnan(y_data) & ~outliers
        y_data_clean = y_data[mask]
        y_pred_clean = y_pred[mask]

@@

        return slope, intercept, r2, corr, outliers
    except Exception:
        return 0,0,0,0,None
```

@@

    # Perform linear comparison between the datasets
    c, m, r2, corr, outliers = compare_data(x, y, sigma)

    # Plot the data with error bars
    # errorbar -> needs np.ma -> no units
    if outliers is not None:
        filtered_data = np.ma.masked_array(y_data, mask = outliers)
        plt.errorbar(x_data, y_data, yerr=yerr_d, c = "black", label="Clipped Data", fmt = "o", fillstyle = "none")
        plt.scatter(x_data, filtered_data,  label="Fitted Data", c = "blue")
    else:
        plt.errorbar(x_data, y_data, xerr = xerr_d, yerr = yerr_d, alpha = 0.3, c = "blue", label = "Data", fmt = ".")

@@

## Luminosities

```{python}
#| echo: false
logKLum_corr = round(compare_data(dt["logKLum"], dt["logL_K"], sigma = True)[3], 3)

@@

```{python}
#| echo: false
scatter_plot(dt["logKLum"], dt["logL_K"], sigma = True)
```

@@

```{python}
#| echo: false
mag_B_corr = round(compare_data(dt["mag_B"], dt["BT"], sigma = True)[3], 3)
Kmag_corr = round(compare_data(dt["Kmag"], dt["K"], sigma = True)[3], 3)
```

@@

```{python}
scatter_plot(dt["mag_B"], dt["BT"], yerr = dt["E_BT"], sigma = True)
scatter_plot(dt["Kmag"], dt["K"], yerr = dt["E_K"], sigma = True)
```

@@
